---
title: "JRSNZ-analysis"
output: html_document
date: "2024-12-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RRST Analysis Script - 
Explorative analysis of RRST pilot data to assess analysis pipeline for interoceptive sensitivity that accounts for physiological variances. 

Data files needed: 
1. ("results-RRST.xlsx") RRST Data 
3.. ("CPQuest.xlsx") Questionnaire Data 
Environment set up is as below: 

```{r Set Up Environment}
rm(list=ls()) # Clear environment

## Load Libraries and Tools
library(ggplot2) # for plotting graphs
library(ggprism) # pretty figures
library(tidyr) # data formatting
library(tidyverse) # data manipulation and visualisation
library(rstatix) # pipe friendly R functions for statistical analysis
library(readxl) # for reading xls files
library(ggpubr) # for plotting QQplot
library(RColorBrewer) # Colour palettes for plots
library(sjPlot) # for plot_model
library(corrplot) # for correlation plots

options(scipen = 999) # to display variable quantities in decimals (not in scientific notation format)
```

Set working directory and load data:
```{r Set working directory and load data}
## SET DIRECTORY
setwd("/Users/ella/Library/Mobile Documents/com~apple~CloudDocs/MSc/MSc-PB/PB-analysisCode/MScPB-analysis")
################################################################################
## LOAD DATA
RRST_df <- read_xlsx("data/results-RRST.xlsx")
QUEST_df <- read_xlsx("data/results-CPQuest.xlsx")
```

Calculate total score for QUEST data, and add new columns for total score:
```{r QUEST Data scoring}
# Calculate total scores for MAIA and assign to new row in QUEST_df
QUEST_df$MAIA <- rowSums(QUEST_df[ , c(2:32), na.rm = TRUE])

# Calculate total scores for ASI and assign to new row in QUEST_df
QUEST_df$ASI <- rowSums(QUEST_df[ , c(33:51), na.rm = TRUE])

# Calculate total scores for PANAST positive Affect score and assign to new row in QUEST_df
QUEST_df$PANAST_P <- rowSums(QUEST_df[ , c(52, 54, 56, 60, 61, 63, 65, 67, 68, 70), na.rm = TRUE])
# Calculate total scores for PANAST negative affect score and assign to new row in QUEST_df
QUEST_df$PANAST_N <- rowSums(QUEST_df[ , c(53, 55, 57, 58, 59, 62, 64, 66, 69, 71), na.rm = TRUE])

# Calculate total scores for PCS_Resp and assign to new row in QUEST_df
QUEST_df$PCS_Resp <- rowSums(QUEST_df[ , c(72:84), na.rm = TRUE])

# Calculate total scores for PVAQ and assign to new row in QUEST_df
QUEST_df$PVAQ_Resp <- rowSums(QUEST_df[ , c(85:100), na.rm = TRUE])

# Calculate total scores for FSS and assign to new row in QUEST_df
QUEST_df$FSS <- rowSums(QUEST_df[ , c(101:108), na.rm = TRUE])

# Calculate total scores for CD_RISC and assign to new row in QUEST_df
QUEST_df$CD_RISC <- rowSums(QUEST_df[ , c(109:132), na.rm = TRUE])

# Calculate total scores for GSE and assign to new row in QUEST_df
QUEST_df$GSE <- rowSums(QUEST_df[ , c(133:142), na.rm = TRUE])

# Calculate total scores for GSE and assign to new row in QUEST_df
QUEST_df$GSE <- rowSums(QUEST_df[ , c(133:142), na.rm = TRUE])
```

Create new dataframe for total scores from QUEST data
```{r QUEST total score df}
QUEST_ts_df <- data.frame(
  PPID = QUEST_df$Participant_ID,
  MAIA = QUEST_df$MAIA,
  ASI = QUEST_df$ASI,
  PANAST_P = QUEST_df$PANAST_P,
  PANAST_N = QUEST_df$PANAST_N,
  PCS_R = QUEST_df$PCS_Resp,
  PVAQ_R = QUEST_df$PVAQ_Resp,
  FSS = QUEST_df$FSS,
  CD_RISC = QUEST_df$CD_RISC,
  GSE = QUEST_df$GSE,
  GAD7 = QUEST_df$GAD7,
  STAI_S = QUEST_df$STAI_S,
  STAI_T = QUEST_df$STAI_T,
  CESD = QUEST_df$CESD)
```

Confirm formatting of imported data:
```{r Format of imported data}
# Check the format of the imported data
head(RRST_df)
head(QUEST_ts_df)
```

Convert data from wide to long format:
```{r Wide to long Df}
# Transform RRST Data into Long Format and add the task column
RRST_long <- RRST_df %>%
  pivot_longer(cols = 2:7, names_to = "metric", values_to = "value") %>%
  mutate(task = ifelse(metric %in% c("Threshold", "Slope", "Confidence", "MetaPerf", "Accuracy", "Intensity"), "RRST"))

# Transform QUEST Data into Long Format, add the task column, 
QUEST_long <-  QUEST_ts_df %>% 
  pivot_longer(cols = 2:14, names_to = "metric", values_to = "value") %>% 
  mutate(task = "QUEST")

# Merge RRST and QUEST Data, then arrange by PPID
df <- bind_rows(RRST_long, QUEST_long) %>%
  arrange(PPID)
```

Collecting/calculating summary stats and testing assumptions: 
Testing for normality of residuals using Shapiro-Wilk (all variables included, excluding slope [as slope is held constant for all participants]); and QQplot for confirmation. QQPlot also used to check assumption of normality. 
 

```{r Summary stats}
brief_stats <- df %>%
  group_by(metric, task) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE),
    count = n()
  ) %>%
  ungroup()

print(brief_stats)

## Test assumption of Normality using Shapiro-Wilk
SW_test <- df %>%
  filter(!grepl("slope", metric, ignore.case = TRUE)) %>%  # Exclude 'slope' metric
  group_by(metric, task) %>%
  summarise(
    shapiro_p_value = shapiro.test(value)$p.value
  ) %>%
  ungroup()

### Create QQPlot for Perceptual Variables
# Filtering df to remove NaN values. 
filtered_df <- df %>%
  filter(metric %in% c("Threshold", "Confidence", "MetaPerf"))%>%
  filter(complete.cases(value)) # Remove NaN values for visualisation

# Create QQ plot
ggqqplot(filtered_df, x = "value", ggtheme = theme_bw(), add = "qqline") +
  facet_wrap(task ~ metric, scales = "free") +
  labs(
    title = "QQ Plot for Threshold, Confidence, and MetaPerf",
    x = "Theoretical Quantiles",
    y = "Sample Quantiles"
  )

### Create QQPlot for Questionnaire Variables
#  Filter data for QUEST
QUEST_QQ_df <- df %>%
  filter(task == "QUEST") %>%
  filter(complete.cases(value)) # Remove NaN values for visualisation

# Create QQ plot
ggqqplot(QUEST_QQ_df, x = "value", ggtheme = theme_bw(), add = "qqline") +
  facet_wrap(task ~ metric, scales = "free") +
  labs(
    title = "QQ Plot for QUEST data",
    x = "Theoretical Quantiles",
    y = "Sample Quantiles"
  )


```

Results from Shapiro-Wilk (a value >0.05 indicates normal distribution) indicate all variables but threshold (for the RRST) and GAD7 are normally distributed. Results from perceptual QQPlot suggest that confidence and metacognitive performance/insight for the RRST are normally distributed, threshold is not normally distributed for RRST. QQplot for questionnaire data indicates all but GAD7 are normally distributed. 

Creating histogram to visualise perceptual variables: 
```{r }
ggplot(filtered_df, aes(x = value, fill = metric)) +
  geom_histogram(position = "identity", bins = 20, alpha = 0.7) +
  facet_grid(task ~ metric, scales = "free") +
  labs(
    title = "Histograms for Threshold, Confidence, and MetaPerf",
    x = "Value",
    y = "Frequency",
    fill = "metric"
  ) +
  theme_bw()
```

```{r Correlation Matrix}
## For correlation matrix wide data frame is required (one row per participant, separate columns for each metric)
corrM_colour <- colorRampPalette(c("navy", "white", "#7AA8C4"))(10) #create colour palette 
## Transforming data to wide format
df_wide <-  QUEST_ts_df %>% 
  left_join(
    RRST_df %>% 
            rename_with(~ paste0(., "_RRST"), -c(PPID)),
            select(PPID, Accuracy_RRST, Slope_RRST, Threshold_RRST, Confidence_RRST, MetaPerf_RRST, Intensity_RRST),
            by = "PPID")
head(df_wide) # to check transform was correct

# Exploring Correlation of RRST slope threshold
correlation_matrix <- cor(df_wide %>% select(-PPID, -Slope_RRST, -Accuracy_RRST), method = "spearman") #create correlation matrix

df_wide_filtered <- df_wide %>% select(-PPID, -Slope_RRST, -Accuracy_RRST) #filter data frame for exact match values (slope), accuracy (as accuracy is held by task algorithms), and character values (PPID)

p_matrix <- cor.mtest(df_wide_filtered , method = "spearman", exact = FALSE) # calculate p value of each Spearman correlation 

# Create Correlation matrix according to specifications in corrplot, including sig level as 
# calculated by cormtest
corrplot(correlation_matrix, p.mat = p_matrix$p, sig.level = c(0.01, 0.05, 0.1), pch.cex = 0.9, insig = 'label_sig', method = "circle", col = corrM_colour, order = 'original', type = "upper", tl.col = "black")

# Correlation matrix with p value added to all correlations
corrplot(correlation_matrix, p.mat = p_matrix$p, insig = 'p-value', sig.level = -1,  method = "circle", col = corrM_colour, order = 'original', type = "upper", tl.col = "black")
```
 
 